### 2016-04-22 manuscript-body.tex
Add citations for knowledge-in-pieces and framing

In 1993, a pair of articles in the learning sciences staked a strong
claim for viewing knowledge as a network of pieces, isolated enough to
be locally triggered but trainable enough to fire in larger concerted
patterns (diSessa, 1993; Smith, diSessa, \& Roschelle, 1993). Informed
in part by agent-based accounts of cognition (Minsky, 1986) and complex
systems models (diSessa, 2002), the central tenets of an ``in-pieces''
approach hold that knowledge is emergent from interacting primitives,
rather than unitary and monolithic. An example from Smith, diSessa, and
Roschelle helps illustrate the point.

@@

In each of the rubberband examples, various pieces of intuitive physical
knowledge describe the mechanism at work: the rubber band binds the
newspaper, grips the jar lid, and acts a source of springiness for the
bobbing object. Although a mapping cannot be made from the rubberband to
scientific entities, it is quite easy to map these qualitatively
distinct physical processes to scientific entities and laws. For
example, instances of binding almost always map to a practically rigid
body. Likewise, gripping maps to friction forces, and springiness maps
to Hooke's law. This suggests that applicability can depend directly on
our intuitive knowledge---knowledge that exists prior to any formal
scientific training (Smith et al., 1993, p. 144).

The in-pieces approach to modeling cognition has been used, among other
things, to explain how experts reason about fractions and decimals
(Smith et al., 1993), how students reason about forces in physics
(diSessa \& Sherin, 1998; diSessa, 1993; Hammer, 1996; Sherin, 2001),
how students construct and evaluate algebraic representations of
physical situations (Izsák, 2004), and how knowledge transfers across
contexts (Hammer et al., 2005; Wagner, 2006). Because its starting
assumption is that knowledge is fragmented, knowledge-in-pieces can
account for wide variations of how people---particularly novices---use
knowledge on a moment-to-moment basis. In other words, because it
assumes knowledge is local, it can still explain the kinds of
globally-inconsistent ways people might reason about physical situations
(diSessa, 1993). As a framework, an in-pieces approach ultimately argues
that models of concept replacement and good/bad criteria for knowledge
should be supplanted by a learning model of alignment/refinement of
prior knowledge and the consideration of knowledge as
productive/unproductive.

Hammer and colleagues have worked to extend the in-pieces approach to
explain how students' epistemological activity---how they orient toward
knowledge and knowing in a context (Hammer et al., 2005; Hammer \& Elby,
2002, 2003). Specifically, those authors use two core theoretical
constructs to explain students' stances toward knowledge and knowing:

\begin{itemize}
\item
  \emph{Epistemological Resources} are the epistemological equivalent of
  diSessa's phenomenological primitives (p-prims). Resources, the
  authors propose, are the atomic units involved in how people cognize
  about the source of knowledge, the nature of knowledge, and
  epistemological activities (Hammer \& Elby, 2003; Louca, Elby, Hammer,
  \& Kagey, 2004).
\item
  \emph{Epistemological Frames} are the emergent result of subsets of
  resources acting in concert. Drawing from both Goffman's (1974)
  sociological notion of frame as structures of expectations and
  subsequent work on framing in discourse (Tannen, 1993),
  epistemological frames are a participant's local answer to the
  question ``what is it {[}specifically, what knowledge activity{]}
  that's going on here'' (Goffman, 1974, p. 8).
\end{itemize}

Resources can frames can interact in activity settings to produce
larger-scale patterns called ``epistemological coherences'' (Rosenberg,
Hammer, \& Phelan, 2006) where evidence from data suggests that a
network of discrete cognitive units can nonetheless give rise to stable
cognition.

An example helps ground this in-pieces approach to epistemology. In
Russ, Coffey, Hammer, and Hutchison (2008), the authors describe the
situation where an elementary student reasons about why an empty juice
box collapses when you suck on the straw. One student gives what the
authors deem to be an excellent mechanistic account of why the juice box
collapses. But, as the teacher seems to steer the discussion toward
vocabulary---in this case, ``pressure''---the student clearly pulls back
from her mechanistic reasoning, seems much more diffident, and claims
that pressure is hard to explain. That example highlights the disconnect
between doing science as knowing vocabulary and doing science as
reasoning mechanistically. Moreover, it strikingly highlights that a
student who by all accounts produced an excellent explanation of how
pressure works was left nonetheless with the impression that pressure
was hard to explain.\footnote{In fairness to the teacher, I'm trying to
  focus on the result of the interaction and far less on the intent the
  teacher might have had. The student still left with a sense that
  science might be about vocabulary, even if that's not the view the
  teacher would espouse or was trying to enact in the moment.}

>>>>

In 1993, a pair of articles in the learning sciences staked a strong
claim for viewing knowledge as a network of pieces, isolated enough to
be locally triggered but trainable enough to fire in larger concerted
patterns \cite{disessa_epistemology_1993,smith_misconceptions_1993}. Informed
in part by agent-based accounts of cognition \cite{minsky_society_1986} and complex
systems models \cite{disessa_why_2002}, the central tenets of an ``in-pieces''
approach hold that knowledge is emergent from interacting primitives,
rather than unitary and monolithic. An example from Smith, diSessa, and
Roschelle helps illustrate the point.

@@

\begin{quote}
  In each of the rubberband examples, various pieces of intuitive physical
  knowledge describe the mechanism at work: the rubber band binds the
  newspaper, grips the jar lid, and acts a source of springiness for the
  bobbing object. Although a mapping cannot be made from the rubberband to
  scientific entities, it is quite easy to map these qualitatively
  distinct physical processes to scientific entities and laws. For
  example, instances of binding almost always map to a practically rigid
  body. Likewise, gripping maps to friction forces, and springiness maps
  to Hooke's law. This suggests that applicability can depend directly on
  our intuitive knowledge---knowledge that exists prior to any formal
  scientific training \cite{smith_misconceptions_1993}.
\end{quote}

The in-pieces approach to modeling cognition has been used, among other
things, to explain how experts reason about fractions and decimals \cite{smith_misconceptions_1993}, how students reason about forces in physics \cite{disessa_what_1998,disessa_epistemology_1993,hammer_misconceptions_1996,sherin_how_2001},
how students construct and evaluate algebraic representations of
physical situations \cite{izsak_we_2003,izsak_students_2004}, and how knowledge transfers across
contexts \cite{hammer_resources_2005,wagner_transfer_2006}. Because its starting
assumption is that knowledge is fragmented, knowledge-in-pieces can
account for wide variations of how people---particularly novices---use
knowledge on a moment-to-moment basis. In other words, because it
assumes knowledge is local, it can still explain the kinds of
globally-inconsistent ways people might reason about physical situations
\cite{disessa_epistemology_1993}. As a framework, an in-pieces approach ultimately argues
that models of concept replacement and good/bad criteria for knowledge
should be supplanted by a learning model of alignment/refinement of
prior knowledge and the consideration of knowledge as
productive/unproductive.

Hammer and colleagues have worked to extend the in-pieces approach to
explain how students' epistemological activity---how they orient toward
knowledge and knowing in a context \cite{hammer_resources_2005,hammer_form_2002,elby_substance_2001}. Specifically, those authors use two core theoretical
constructs to explain students' stances toward knowledge and knowing:

\begin{itemize}
\item
  \emph{Epistemological Resources} \cite{hammer_resources_2005} are the epistemological equivalent of
  diSessa's phenomenological primitives (p-prims) \cite{disessa_epistemology_1993}. Resources, the
  authors propose, are the atomic units involved in how people cognize
  about the source of knowledge, the nature of knowledge, and
  epistemological activities \cite{hammer_tapping_2003,louca_epistemological_2004}.
\item
  \emph{Epistemological Frames} are the emergent result of subsets of
  resources acting in concert. Drawing from both Goffman's
  sociological notion of frame as structures of expectations \cite{goffman_frame_1974} and
  subsequent work on framing in discourse \cite{tannen_framing_1993},
  epistemological frames are a participant's local answer to the
  question ``what is it {[}specifically, what knowledge activity{]}
  that's going on here'' \cite{goffman_frame_1974}.
\end{itemize}

Resources can frames can interact in activity settings to produce
larger-scale patterns called ``epistemological coherences'' \cite{rosenberg_multiple_2006} where evidence from data suggests that a
network of discrete cognitive units can nonetheless give rise to stable
cognition.

An example helps ground this in-pieces approach to epistemology. Russ, Coffey, Hammer, and Hutchison, describe the
situation where an elementary student reasons about why an empty juice
box collapses when you suck on the straw \cite{russ_making_2008}. One student gives what the
authors deem to be an excellent mechanistic account of why the juice box
collapses. But, as the teacher seems to steer the discussion toward
vocabulary---in this case, ``pressure''---the student clearly pulls back
from her mechanistic reasoning, seems much more diffident, and claims
that pressure is hard to explain. That example highlights the disconnect
between doing science as knowing vocabulary and doing science as
reasoning mechanistically. Moreover, it strikingly highlights that a
student who by all accounts produced an excellent explanation of how
pressure works was left nonetheless with the impression that pressure
was hard to explain.\footnote{In fairness to the teacher, I'm trying to
  focus on the result of the interaction and far less on the intent the
  teacher might have had. The student still left with a sense that
  science might be about vocabulary, even if that's not the view the
  teacher would espouse or was trying to enact in the moment.}
